---
title: ""
author: "Shane Kercheval"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github('shane-kercheval/rtools')
library(rtools)
# library(scales)
# library(stringr)
library(lubridate)
# library(ggrepel)
# library(forecast)
library(tidyverse)
library(ggplot2)
library(knitr)

options(scipen=999) # non-scientific notation
options(dplyr.summarise.inform=F)

source('../regression/regression_helpers.R')

theme_set(theme_light())

calculate_plot_width <- function(plot_height) { plot_height * 1.61803398875 }
plot_width_height_6 <- calculate_plot_width(6)
plot_width_height_7 <- calculate_plot_width(7)
plot_width_height_8 <- calculate_plot_width(8)
```

> A diff-in-diff analysis applies when you have two groups whose pretreatment differences can be isolated and modeled so that post-treatment differences are the basis for causal estimates of the treatment effect. - Business Data Science pg 141

Author's Original Code: https://github.com/TaddyLab/BDS/blob/master/examples/paidsearch.R

See my questions: https://github.com/TaddyLab/BDS/issues/17

```{r}
# business data science pg 141
sem <- read.csv("data/paidsearch.csv")
sem <- sem %>%
    mutate(date = as.Date(strptime(date, "%d-%b-%y", tz = 'UTC')),
           dma = factor(dma),
           variant = ifelse(search.stays.on == 1, 'Control', 'Treatment')) %>%
    relocate(variant, .after=search.stays.on)
head(sem)
```

Here is the overall data, because the treatment is much smaller in total revenue, it is difficult to see if there is a noticable drop after Adwords was shut off.

```{r message=FALSE, warning=FALSE}
sem %>%
    count(date, variant, wt=revenue, name='revenue') %>%
    ggplot(aes(x=date, y=revenue, color=variant, group=variant)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth() +
    labs(title="Daily Revenue",
         subtitle="The 'Control' are the DMAs where adwords will be left on.\nThe 'Treatment' are the DMAs where adwords is turned off.\nWe could expect a drop in the Treatment as compared to the Control.")
```

Using log() makes it easier to visualize.

```{r message=FALSE, warning=FALSE}
sem %>%
    count(date, variant, wt=revenue, name='revenue') %>%
    ggplot(aes(x=date, y=log(revenue), color=variant, group=variant)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth()
```

> Here we will focus on differences in `log` revenue because we anticipate that the treatment and control groups are related on percentage scale. BDS pg 143.

```{r message=FALSE, warning=FALSE}
sem %>%
    group_by(date) %>%
    summarise(difference = log(mean(revenue[variant == 'Treatment'])) - log(mean(revenue[variant == 'Control']))) %>%
    ungroup() %>%
    ggplot(aes(x=date, y=difference)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth() +
    labs(title = "Log Difference of Average Revenue (Treatment - Control)",
         subtitle = "Decrease indicates a drop in the Treatment compared to the Control",
         y = "Log Difference (Treatment - Control)",
         x = 'Date')
```

(Sum would give us the same pattern/trend on different scale)

```{r message=FALSE, warning=FALSE}
sem %>%
    group_by(date) %>%
    summarise(difference = log(sum(revenue[variant == 'Treatment'])) - log(sum(revenue[variant == 'Control']))) %>%
    ungroup() %>%
    ggplot(aes(x=date, y=difference)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth() +
    labs(title = "Log Difference of Total Revenue (Treatment - Control)",
         subtitle = "Decrease indicates a drop in the Treatment compared to the Control",
         y = "Log Difference (Treatment - Control)",
         x = 'Date')
```

Same but using separate `lm` on before vs. after.

```{r message=FALSE, warning=FALSE}
sem %>%
    group_by(date, treatment_period) %>%
    summarise(difference = log(mean(revenue[variant == 'Treatment'])) - log(mean(revenue[variant == 'Control']))) %>%
    ungroup() %>%
    ggplot(aes(x=date, y=difference)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth(method='lm', aes(group=treatment_period)) +
    labs(title = "Log Difference of Average Average (Treatment - Control)",
         subtitle = "Decrease indicates a drop in the Treatment compared to the Control",
         y = "Log Difference (Treatment - Control)",
         x = 'Date')
```

## Quick Refresher & Example with Logs

If we would not have used log differences, here is what we would have had. The difference is dominated by the larger group (control).

```{r message=FALSE, warning=FALSE}
sem %>%
    group_by(date, treatment_period) %>%
    summarise(difference = mean(revenue[variant == 'Treatment']) - mean(revenue[variant == 'Control'])) %>%
    ungroup() %>%
    #pivot_longer(-date, names_to=)
    ggplot(aes(x=date, y=difference)) +
    geom_vline(xintercept = ymd('2012-05-22')) +
    geom_line() +
    scale_x_date() +
    scale_y_continuous(labels=rt_pretty_axes) +
    geom_smooth(method='lm', aes(group=treatment_period)) +
    labs(title = "Difference of Average Revenue (Treatment - Control)",
         subtitle = "Note: as compared with log difference, no change in trend is found,\nbecause we are comparing a change within a smaller group to no change in a larger group\n(i.e. additive difference), whereas we want to know the multiplicative difference.",
         y = "Difference (Treatment - Control)",
         x = 'Date')
```

For example, `.y` and `.z` change multiplicatively (actually percent change is the same for each, which you can see in the last graph)

```{r}
.x <- seq(1, 2, .1)
.x <- c(.x, rev(.x))
(.y <- .x * 1000)
(.z <- .y * 1000)
```

Both sequences above have the same percent change.

However, there additive difference doesn't reflect that.

```{r}
.z - .y
```

But when we look at the log differences, it is exactly the same for each value, because each variable is changing at constant rate, and so there log differences are the same... i.e. they change multiplicatively.

```{r}
log(.z) - log(.y)
```

Both differences are graphed below.

```{r}
data.frame(index=1:length(.y), y=.y, z=.z) %>%
    pivot_longer(cols=-index) %>%
    ggplot(aes(x=index, y=value, color=name, group=name)) +
    geom_line()
```

```{r}
data.frame(index=1:length(.y), y=.y, z=.z) %>%
    mutate(differences = z - y) %>%
    ggplot(aes(x=index, y=differences)) +
    geom_line()
```

```{r}
data.frame(index=1:length(.y), y=.y, z=.z) %>%
    mutate(difference = log(z) - log(y)) %>%
    ggplot(aes(x=index, y=difference)) +
    geom_line() +
    labs(y='log differences', 
         subtitle = 'both .x and .y are increasing and decreasing at the same rate')
```

## Create dataset that is basis for regression model.

Note: the author uses `mean(log(revenue))` but seems like we would want to use `log(mean(revenue))`.

This is especially true if we were to use `sum` rather than mean since control group is much larger and we would want to take the log of the total (like we did in the graphs).

A model using `log(mean(revenue))` vs `log(sum(revenue))` gives the same interaction coefficient, which is not true for `mean(log(revenue))` vs `sum(log(revenue))`.

I ended up using `log(sum(revenue))`, which again is the same as `log(mean(revenue))` which gives an interaction coefficient of `-0.005775` compared to `mean(log(revenue))` which gives `-0.006587` which matches the author's results.

```{r}
# treatment_period of 0 means < May 22, when the ads were on for both variants (consistance difference is expected)
# treatment_period of 1 means >= May 22, when the ads were shut off for Treatment (potental drop in sales for treatment may be expected)
sem_per_dma <- sem %>%
    group_by(dma, treatment_period, variant) %>%
    summarise(log_total_revenue = log(sum(revenue)), n=n()) %>%
    ungroup() %>%
    mutate(variant = factor(variant, levels=c("Control", "Treatment"))) %>%
    as.data.frame()
head(sem_per_dma)
```

## Create Model

```{r}
reg_results <- lm(log_total_revenue ~ treatment_period * variant, data=sem_per_dma)
summary(reg_results)
```

```{r}
get_regression_equation(reg_results, .round_by = 5)
```

> This says that the treatment - turning off paid search ads - has a small but not statistically significant negative effect on log revenue (0.66 percentage point drop with a standard error of 0.55% points). (pg 145.)

This is a `log-level` model so pg 39 of Intro to econometrics says that `1 unit` change in x has a `B * 100` percent change in `y`.

So, `-0.006587 * 100` == `-0.6587` i.e. `0.66%` drop.

```{r}
(interaction_coeff <- coefficients(reg_results)['treatment_period:variantTreatment'])
(interaction_coeff <- paste0(round(interaction_coeff * 100, 2), '%'))
```

## Coefficient interpretation

Intercept coefficient is the average `log(sum(revenue))` of control before applying treatment.

```{r}
sem_per_dma %>%
    filter(variant == 'Control' & treatment_period == 0) %>%
    pull(log_total_revenue) %>%
    mean()
 coefficients(reg_results)['(Intercept)']
```

The coefficient for `treatment_period` is the difference in average `log(sum(revenue))` before/after for the `control`

```{r}
sem_per_dma %>%
    filter(variant == 'Control') %>%
    group_by(treatment_period) %>%
    summarise(mean_y = mean(log_total_revenue)) %>%
    ungroup() %>%
    summarise(diff = mean_y[treatment_period == 1] - mean_y[treatment_period == 0]) %>%
    pull(diff)
 coefficients(reg_results)['treatment_period']
```

The coefficient for `variantTreatment` is the difference in average `log(sum(revenue))` between the control & variant during the treatment_period of 0.

```{r}
sem_per_dma %>%
    filter(treatment_period == 0) %>%
    group_by(variant) %>%
    summarise(mean_y = mean(log_total_revenue)) %>%
    ungroup() %>%
    summarise(diff = mean_y[variant == 'Treatment'] - mean_y[variant == 'Control']) %>%
    pull(diff)
 coefficients(reg_results)['variantTreatment']
```

The coefficient for `treatment_period:variantTreatment`, is the **difference over time** (i.e. period 1 vs period 2) in the **difference of average** `log(sum(revenue))` in the two groups (treatment vs control)... i.e. **difference in difference**.

```{r}
sem_per_dma %>%
    group_by(treatment_period, variant) %>%
    summarise(y = mean(log_total_revenue)) %>%
    ungroup() %>%
    summarise(diff = (y[variant == 'Treatment' & treatment_period == 1] -
                      y[variant == 'Control' & treatment_period == 1]) - 
                     (y[variant == 'Treatment' & treatment_period == 0] -
                      y[variant == 'Control' & treatment_period == 0]))
```

```{r}
coefficients(reg_results)['treatment_period:variantTreatment']
```

## Controlling for DMA-sepciic revenue levels

```{r}
summary(reg_results)
```

> Note that these rows of data do not satisfy the standard independence assumption: each two observations on the same DMA will be correlated. ... you can use clustered starndard error, or you can get a similar result by controlling for the DMA-specific revenue levels in the regression equation. BDS pg 144-145

```{r}
reg_results <- lm(log_total_revenue ~ dma + treatment_period * variant, data=sem_per_dma)
reg_results %>%
    tidy() %>%
    filter(!str_detect(term, 'dma'))
```
