---
title: ""
author: "Shane Kercheval"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github('shane-kercheval/rtools')
#library(rtools)
# library(scales)
# library(stringr)
# library(lubridate)
# library(ggrepel)
# library(forecast)
library(tidyverse)
library(ggplot2)
library(knitr)

options(scipen=999) # non-scientific notation
options(dplyr.summarise.inform=F)

theme_set(theme_light())

calculate_plot_width <- function(plot_height) { plot_height * 1.61803398875 }
plot_width_height_6 <- calculate_plot_width(6)
plot_width_height_7 <- calculate_plot_width(7)
plot_width_height_8 <- calculate_plot_width(8)
```

# Packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(ggplot2)
library(tidytext)
```

# Data Cleaning

---

```{r eval=FALSE, include=FALSE}
CO2
starwars
storms
```
## General

### `clean_names()`

```{r}
iris %>% colnames()
```

```{r}
iris %>% janitor::clean_names() %>% colnames()
```

---

### `extract()`

Turn `1` column into `x` columns based on regex

`convert=TRUE` converts them to numeric

```{r}
.df <- data.frame(season_epison = paste0('S', c(1, 1, 1, 2, 2, 2, 3, 3, 3),
                                         'E', c(1, 2, 3, 1, 2, 3, 1, 2, 3)))

.df %>% extract(season_epison, c('season', 'episode'), 'S(.*)E(.*)', convert = TRUE, remove = FALSE)
```

---


## Aggregation

### `group_by() & which.max()`

```{r}
which.max(c(2, 1, 4, 3))
which.max(c(2, 4, 4, 3))
```

`first(name[which.max(height)])`

```{r}
starwars %>%
    group_by(gender) %>%
    summarise(n = n(),
              tallest_person = first(name[which.max(height)]),
              tallest_height = max(height, na.rm = TRUE),
              oldest_person = first(name[which.min(birth_year)]))
```

[Tidy Tuesday screencast: analyzing franchise revenue - YouTube](https://youtu.be/1xsbTs9-a50?t=365)

---

### `group_by()` & `top_n()`

This gets the `N` rows associated with the top `N` values for each category being grouped

`top_n(3, height)`

```{r}
starwars %>%
    group_by(gender) %>%
    top_n(3, height) %>%
    select(gender, name, height) %>%
    arrange(gender, height) %>%
    ungroup()
```

---

### `summarise()` & `across`

* Summarize multiple columns with multiple functions
* name columns with `glue` style convention

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(across(starts_with("d"),
                   list(mean = mean,
						   sd = sd),
                   .names = "{col}_{fn}"))
```

---

same as above using formulas

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarise(across(starts_with("d"), 
                   list(minus_sd = ~ (mean(.x) - sd(.x)), 
                        mean = mean, 
                        plus_sd = ~ (mean(.x) + sd(.x)))
                   ))
```

---

### `summarise_at()`

```{r}
iris %>% group_by(Species) %>% summarise_at(vars(Sepal.Length, Sepal.Width), sum, na.rm = TRUE)
```

---

### `add_count`

```{r}
iris %>% add_count(Species, name = 'num_species') %>% head()
```

`add_count(Species, name = 'num_species')` is equivalent to `group_by(Species) %>% mutate(num_species = n()) %>% ungroup()`

---

# dplyr/tidyverse 

## Indirection

Create a dplyr-like function that uses the column names of the dataframe rather than strings or the object directly.

https://dplyr.tidyverse.org/articles/programming.html

### dplyr-like function

```{r}
var_summary <- function(data, var) {
    # note the {{ var }}
    data %>%
        summarise(n = n(),
                  min = min({{ var }}),
                  max = max({{ var }}))
}

mtcars %>% 
    group_by(cyl) %>% 
    var_summary(mpg)
```

> If you want the user to provide a set of data-variables that are then transformed, use across():

```{r}
var_summary <- function(data, .group_by, .var) {
    # note the {{ var }}
    data %>%
        group_by(across({{ .group_by }})) %>%
        summarise(n = n(),
                  min = min({{ .var }}),
                  max = max({{ .var }}))
}

mtcars %>% var_summary(.group_by=cyl, .var=mpg)
```

```{r}
mtcars %>% var_summary(.group_by=c(cyl, vs), .var=mpg)
```

---

> Use the .names argument to across() to control the names of the output.

(Note the vignette uses `.col` instead of `col`, which fails.)

```{r}
my_summarise <- function(data, .group_by, .summarise_vars) {
  data %>%
    group_by(across({{ .group_by }})) %>% 
    summarise(across({{ .summarise_vars }},
                     mean,
                     .names = "mean_{col}"))
}
mtcars %>% my_summarise(cyl, mpg)
```

```{r}
mtcars %>% my_summarise(c(cyl, vs), .summarise_vars=c(mpg, hp))
```

---

### .data

> "Note that `.data` is not a data frame; it’s a special construct, a pronoun, that allows you to access the current variables either directly, with `.data$x` or indirectly with `.data[[var]]`. Don’t expect other functions to work with it."

```{r}
var <- 'cyl'
mtcars %>% count(.data[[var]])
```

```{r}
# ROW_NUMBER() OVER (PARITION BY col_x ORDER BY col_y) AS index
row_number_over_partition_by <- function(data, .partition_by, .order_by) {
    data %>%
        group_by(across({{ .partition_by }})) %>%
        mutate(index = row_number({{ .order_by }})) %>%
        ungroup()
}
mtcars %>% 
    select(cyl, mpg) %>%
    row_number_over_partition_by(cyl, mpg) %>%
    arrange(cyl, mpg)
```

```{r}
mtcars %>% 
    select(cyl, mpg) %>%
    row_number_over_partition_by(cyl, desc(mpg)) %>%
    arrange(cyl, mpg)
```

### eval_tidy

```{r}
with_data <- function(data, .x) {
  expr <- rlang::enquo(.x)
  print(expr)
  rlang::eval_tidy(expr, data = data)
}
mtcars %>% with_data(.x=mean(cyl) * 10)
```

### other examples

```{r}
# this is a hack because row_number only takes one column
#.partition_by, 
row_number_over_partition_by <- function(.x, .col_name, .partition_by, ...) {
    .result <- .x %>%
        mutate(temp______original_order = row_number()) %>%
        group_by(across({{ .partition_by }})) %>%
        arrange( ... , .by_group = TRUE) %>%
        mutate(grouped_index = row_number()) %>%
        ungroup() %>%
        arrange(temp______original_order) %>%
        pull(grouped_index)

    .x[deparse(substitute(.col_name))] <- .result
    
    return (.x)
}
mtcars %>% 
    select(cyl, am, vs, mpg) %>%
    row_number_over_partition_by(
        # new column name
        .col_name = my_index,
        # "partition by"
        .partition_by = c(cyl, am),
        # "order by" i.e. sort index in order of these variable
        desc(vs), mpg
    ) %>%
    arrange(desc(cyl), am, desc(vs), mpg)
```

```{r}
filter_over_partition_by <- function(.x, .partition_by, ...) {
    .x %>%
        mutate(temp______original_order = row_number()) %>%
        group_by(across({{ .partition_by }})) %>%
        arrange( ... , .by_group = TRUE) %>%
        filter(row_number() == 1) %>%
        ungroup() %>%
        arrange(temp______original_order) %>%
        select(-temp______original_order)
}
mtcars %>% 
    select(cyl, am, vs, mpg) %>%
    filter_over_partition_by(
        # "partition by"
        .partition_by = c(cyl, am),
        # "order by" i.e. sort index in order of these variable
        desc(vs), mpg
    ) %>%
    arrange(desc(cyl), am, desc(vs), mpg)
```


# ggplot

### `reorder_within()`

Reorder sub-categories within category e.g. for faceting.

For example, the order of `setosa`, `versicolor`, `virginica` is allowed to change for each measurement (e.g. `Petal.Length`, `Sepal.Width`)

```{r}
iris_gathered <- iris %>% pivot_longer(-Species, names_to = 'metric', values_to = 'value')
iris_gathered %>%
    ggplot(aes(x=reorder_within(x=Species, by=value, within=metric),
               y=value)) +
    geom_boxplot() +
    # this is necessary to undo the transformations to the values that reorder_within did
    scale_x_reordered() +
    facet_wrap(~ metric, scales = 'free_x')
```

### `scale_x_log10()` with seconds

	* scales the x-axis ticks but keeps the actual values the same
	* this works when viewing time (in seconds)

```{r}
data.frame(value = abs(rnorm(1000, sd = (60^4) / 3))) %>%
ggplot(aes(x='', y=value)) +
    geom_point() +
    scale_y_log10(breaks = c(0, 60 ^ (0:4)),
                  labels = c("0", "Second", "Minute", "Hour", "2.5 Days", "120 Days"))
```
