---
title: "Statistical Rethinking"
author: "Shane Kercheval"
output:
  md_document:
    variant: markdown_github
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
#devtools::install_github('shane-kercheval/rtools')
#library(rtools)
#library(stringr)
#library(ggrepel)
#library(forecast)
#library(scales)
#library(lubridate)

library(knitr)

calculate_plot_width <- function(plot_height) { plot_height * 1.61803398875 }
plot_width_height_5 <- calculate_plot_width(5)
plot_width_height_6 <- calculate_plot_width(6)
plot_width_height_7 <- calculate_plot_width(7)
plot_width_height_8 <- calculate_plot_width(8)

pivot_longer_all <- function(.x, names_to='name', values_to='value') {
    .x %>%
        mutate(temp____index = row_number()) %>%
        pivot_longer(-temp____index, names_to=names_to, values_to=values_to) %>%
        dplyr::select(-temp____index)
}
```

# Overview

This document includes examples and exercises from `Statistical Rethinking, Second Edition` by Richard McElreath.

This is a sandbox to experiment with the concepts. Most of the code is from the book (http://xcelab.net/rmpubs/sr2/code.txt) and the links below that provide additional references.

# Resources

- [Book website](http://xcelab.net/rm/statistical-rethinking/)
- [Course Slides and Videos](https://github.com/rmcelreath/stat_rethinking_2020#calendar--topical-outline)
    - [2019](https://github.com/rmcelreath/statrethinking_winter2019)
- [Code](http://xcelab.net/rmpubs/sr2/code.txt)
- [Homework Solutions](https://github.com/rmcelreath/stat_rethinking_2020/tree/main/homework)
    - [2019](https://github.com/rmcelreath/statrethinking_winter2019/tree/master/homework)
- [Stan & Tidyverse](https://vincentarelbundock.github.io/rethinking2/)
- [brms & Tidier](https://bookdown.org/content/4857)
    - [brms](https://github.com/paul-buerkner/brms)
- [Python](https://github.com/pymc-devs/resources/tree/master/Rethinking_2)

# Packages

## statistical packages

```{r message=FALSE, warning=FALSE}
# install.packages(c('coda', 'mvtnorm', 'devtools', 'dagitty'))
# library(devtools)
# devtools::install_github('rmcelreath/rethinking')
library(rethinking)
library(rstan)
library(rstanarm)
#install.packages('brms')
library(brms)
library(bayesplot)
library(loo)
```

## base packages

```{r message=FALSE, warning=FALSE}
library(broom)
library(broom.mixed)

library(tidyverse)
library(ggplot2)
library(scales)
library(ggridges)
```

## Settings

```{r}
theme_set(theme_light())
options(scipen=999) # non-scientific notation
options(dplyr.summarise.inform=F)
```

# Chapter 2

## Bayesian Updating

dbinom gives the probability of getting `x` "successes" given a sample size of `size` and a certain probability.

So `dbinom(x=5, size=10, prob=0.5)`

```{r}
dbinom(x=6, size=9, prob=0.5)
```

so if we say the probability of a "successful event" is 50% (e.g. flipping a coin and landing on heads) and that we are going to flip the coin 10 times (i.e. sample size of 10) the probability we will get exactly 5 heads (i.e. successes) is 24.6%. 

```{r}
plot(x=0:10, y=dbinom(x=0:10, size=10, prob=0.5), type='h')
```

```{r}
sum(dbinom(x=0:10, size=10, prob=0.5))
```

But we don't want the probability distrubtion of seeing heads (i.e. assuming a constance 50% prob of flipping heads), which is what the graph above shows. 

We want the **relative** probability/plausibility of seeing six heads out of 9 flips, assuming all sorts of probablities of flipping heads.

Or to tie into the example, we want the **relative** probability/plausibility of seeing six W's heads out of globe tosses, at each possible value of `p`, the proportion of water, which takes on possible values of `0`-`1` (i.e. no water to all water.).

So if we have `dbinom(x=6, size=9, prob = X)` over a range of probabilities, as we do below, the area under the curve will not sum to 1. This is not a probability distribution it is a distribution of relative plausibilities. 

```{r}
most_plausible <- optimize(function (.x) -dbinom(x=6, size=9, prob = .x),
                           c(0, 1),
                           tol = 0.0001)
most_plausible
```

```{r}
dbinom(x=6, size=9, prob = most_plausible$minimum)
sum(dbinom(0:2, 2, 0.5))
```

```{r}
curve( dbinom(x=2, size=3, prob=x), 0, 1, lty=3, col='gray', xlab = '', ylab='')
curve( dbinom(x=5, size=8, prob=x), 0, 1, add=T, lty=2, xlab = '', ylab='')
curve( dbinom(x=6, size=9, prob=x), 0, 1, add=T, col='blue', xlab = '', ylab='')
abline(v=most_plausible$minimum, col="red", lty=2, lwd=3)
title(xlab="Probability of Water", ylab="dbinom(x=6, size=9, prob=x) (plausibility)")
```

We can standardize the relative plausibilities (based on the fact that curve plots 101 points) by calculating dbinom for a given x, size, and prob, and summing the dbinom for the same x & size over 101 points points between 0 and 1.

The `sum(dbinom(x=.x, size=.size, prob=seq(0, 1, length.out = .length.out)` part is the `P(W, L)` i.e. `average probability of the data` the thing the standardizes the posterior.

Now the values will add up to 1

```{r}
dbinom_density_to_probability <- function(.x, .size, .prob, .length.out=101) {
    dbinom(x=.x, size=.size, prob=.prob) / sum(dbinom(x=.x, size=.size, prob=seq(0, 1, length.out = .length.out)))
}
sum(dbinom_density_to_probability(.x=6, .size=9, .prob = seq(0, 1, length.out = 101)))
```

We see that more information now makes the density higher. The graph is now consistent with the bottom right fiture on page 30.

```{r}
curve( dbinom_density_to_probability(.x=6, .size=9, .prob = x), 0, 1, lwd=2, col='blue', xlab = '', ylab='')
curve( dbinom_density_to_probability(.x=5, .size=8, .prob = x), 0, 1, add=T, lty=2, xlab = '', ylab='')
curve( dbinom_density_to_probability(.x=2, .size=3, .prob = x), 0, 1, add=T, lty=2, col='gray', xlab = '', ylab='')
abline(v=most_plausible$minimum, col="red", lty=2, lwd=3)
title(xlab="Probability of Water", ylab="dbinom(x=6, size=9, prob=x) (posterior probablity)")
```

```{r include=FALSE}
rm(most_plausible)
rm(dbinom_density_to_probability)
```

from https://bookdown.org/content/4857/small-worlds-and-large-worlds.html#building-a-model

```{r}
d <- tibble(toss = c("w", "l", "w", "w", "w", "l", "w", "l", "w")) %>%
    mutate(n_trials  = 1:9,
           n_success = cumsum(toss == "w"))
d
```

```{r}
sequence_length <- 50
d <- d %>% 
  expand(nesting(n_trials, toss, n_success), 
         p_water = seq(from = 0, to = 1, length.out = sequence_length)) %>% 
  group_by(p_water) %>% 
 # you can learn more about lagging here: https://www.rdocumentation.org/packages/stats/versions/3.5.1/topics/lag or here: https://dplyr.tidyverse.org/reference/lead-lag.html
  mutate(lagged_n_trials  = lag(n_trials, k = 1),
         lagged_n_success = lag(n_success, k = 1)) %>% 
  ungroup() %>% 
  mutate(prior      = ifelse(n_trials == 1, .5,
                             dbinom(x    = lagged_n_success, 
                                    size = lagged_n_trials, 
                                    prob = p_water)),
         likelihood = dbinom(x    = n_success, 
                             size = n_trials, 
                             prob = p_water),
         strip      = str_c("n = ", n_trials)) %>% 
  # the next three lines allow us to normalize the prior and the likelihood, 
  # putting them both in a probability metric 
  group_by(n_trials) %>% 
  mutate(prior      = prior / sum(prior),
         likelihood = likelihood / sum(likelihood))

head(d)
```

```{r}
  # plot!
d %>%
  ggplot(aes(x = p_water)) +
  geom_line(aes(y = prior), 
            linetype = 2) +
  geom_line(aes(y = likelihood)) +
  scale_x_continuous("proportion water", breaks = c(0, .5, 1)) +
  scale_y_continuous("plausibility", breaks = NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~strip, scales = "free_y")
```

```{r include=FALSE}
rm(d)
```

## Grid Approximation

For each possible value of `p`, the proportion of water, what is the relatively plausibility of observing the data (6 Ws in 9 tosses) for that value of p.

```{r}
## R code 2.3
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )
# define prior
prior <- rep( 1 , 20 )
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
```

```{r}
prior
likelihood
sum(likelihood)  # doesn't sum to 1, not a probability distribution
```

```{r}
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
```

```{r}
sum(unstd.posterior)
sum(posterior)
```

```{r}
posterior
```

```{r}
## R code 2.4
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```


Or, https://bookdown.org/content/4857/small-worlds-and-large-worlds.html#building-a-model

```{r}
d <- tibble(p_grid = seq(from = 0, to = 1, length.out = 20),      # define grid
            prior  = 1) %>%                                       # define prior
    mutate(likelihood = dbinom(6, size = 9, prob = p_grid)) %>%   # compute likelihood at each value in grid
    mutate(unstd_posterior = likelihood * prior) %>%              # compute product of likelihood and prior
    mutate(posterior = unstd_posterior / sum(unstd_posterior))    # standardize the posterior, so it sums to 1
d
```

```{r}
d %>% 
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_point() +
  geom_line()
```

Now same thing with a different prior.

```{r}
prior <- ifelse( p_grid < 0.5 , 0 , 1 )  # makes anything less than 0.5 have 0 plausibility.
#prior <- exp( -5*abs( p_grid - 0.5 ) )
```

```{r}
## R code 2.3
# define grid
p_grid <- seq( from=0 , to=1 , length.out=20 )
# compute likelihood at each value in grid
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
# compute product of likelihood and prior
unstd.posterior <- likelihood * prior
# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```

```{r include=FALSE}
rm(p_grid)
rm(prior)
rm(likelihood)
rm(unstd.posterior)
rm(posterior)
rm(sequence_length)
rm(d)
```

## Quadratic Approximation (`quap`)

Same thing from above condensed. Note that we pass the same functions to `quap`

```{r}
p_grid <- seq(0, 1, 0.01)
prior <- dunif(p_grid, 0,1)
likelihood <- dbinom(6, 9, p_grid)
posterior <- prior * likelihood / sum(prior * likelihood)
sum(posterior)
```

```{r}
plot(seq(0, 1, 0.01), dbinom(6, 9 , seq(0, 1, 0.01)) * dunif(seq(0,1,0.01), 0,1) / sum(dbinom(6, 9 , seq(0, 1, 0.01)) * dunif(seq(0,1,0.01), 0,1)))
```

```{r}
globe.qa <- quap(
    alist(
        # X ~ B(n, p), that is, X is a binomially distributed random variable, n being the total number of experiments and p the probability of each experiment yielding a successful result
        W ~ dbinom( W+L , p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(W=6,L=3) 
)

# display summary of quadratic approximation
precis( globe.qa , digits = 4)
```

```{r}
globe.qa <- quap(
    alist(
        W ~ dbinom( W+L , p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(W=12,L=6) 
)

# display summary of quadratic approximation
precis( globe.qa , digits = 4)
```

```{r}
y <- c(1, 1, 1, 1, 1, 1, 0, 0, 0)
summary(lm(y~1))
```

```{r}
set.seed(1)
print(rstanarm::stan_glm(y~1, refresh = 0), 3)
```

```{r}
prop.test(x=6, n=9, conf.level = 0.89)
```

```{r include=FALSE}
rm(p_grid)
rm(prior)
rm(likelihood)
rm(posterior)
rm(globe.qa)
rm(y)
```

## MCMC

```{r}
## R code 2.8
n_samples <- 1000
p <- rep( NA , n_samples )
p[1] <- 0.5
W <- 6
L <- 3
for ( i in 2:n_samples ) {
    p_new <- rnorm( 1 , p[i-1] , 0.1 )
    if ( p_new < 0 ) p_new <- abs( p_new )
    if ( p_new > 1 ) p_new <- 2 - p_new
    q0 <- dbinom( W , W+L , p[i-1] )
    q1 <- dbinom( W , W+L , p_new )
    p[i] <- ifelse( runif(1) < q1/q0 , p_new , p[i-1] )
}

## R code 2.9
dens( p , xlim=c(0,1) )
curve( dbeta( x , W+1 , L+1 ) , lty=2 , add=TRUE )
```

```{r include=FALSE}
rm(n_samples)
rm(p)
rm(W)
rm(L)
rm(p_new)
rm(q0)
rm(q1)
rm(i)
```

---

## Homework

https://github.com/rmcelreath/stat_rethinking_2020/blob/main/homework/week01/week01_solutions.pdf

1. Suppose the globe tossing data (Chapter 2) had turned out to be 4 water in 15 tosses. Construct the posterior distribution, using grid approximation. Use the same flat prior as in the book.

```{r}
# remember p_grid is the possible values of the proprortion of water
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
prob_data <- dbinom( 4 , size=15 , prob=p_grid )
posterior <- prob_data * prior
posterior <- posterior / sum(posterior)
plot(p_grid, posterior)
```

```{r}
sum(posterior)
```

> When you plot the result, the posterior will have much more mass over values below 0.5, since that is what the sample indicates. The posterior mean is
about 0.30.

```{r}
set.seed(1)
# p_grid is the possible values of the proprortion of water
# posterior is the probability of that proportion being correct
# we are going to sample from the possible values, relative to the probablity of each value
samples <- sample( p_grid , prob=posterior , size=1e4, replace=TRUE )
mean(samples)
sd(samples)
```

---

2. Start over in 1, but now use a prior that is zero below p = 0.5 and a constant above p = 0.5. This corresponds to prior information that a majority of the Earth’s surface is water. What difference does the better prior make?

```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- c( rep( 0 , 500 ) , rep( 1 , 500 ) )
prob_data <- dbinom( 4 , size=15 , prob=p_grid )
posterior <- prob_data * prior
posterior <- posterior / sum(posterior)
plot(p_grid, posterior)
```

```{r}
set.seed(1)
# bootstrap
samples <- sample( p_grid , prob=posterior , size=1e4, replace=TRUE )
mean(samples)
sd(samples)
```

---

3. For the posterior distribution from 2, compute 89% percentile and HPDI intervals. Compare the widths of these intervals. Which is wider? Why? If you had only the information in the interval, what might you misunderstand about the shape of the posterior distribution?

```{r}
quantile(samples, c(0.055, 0.945))
```

```{r}
set.seed(100)
PI( samples )
HPDI( samples )
```

> The percentile interval (the top one) is wider. The lower bound is above
0.5 (a little bit) and the upper bound is 0.63. The HPDI (bottom interval)
is narrower. The lower bound is right at 0.50, but the upper bound is 0.60
instead of 0.63. You should expect the HPDI to be narrower and the include
the point with highest posterior probability. This is just like the example
on page 57 in the book. But since the boundaries of these intervals aren’t
really informative—nothing special happens at the boundary—when these
intervals are very different, the best thing is not to report intervals at all. Just
draw the posterior distribution, so your colleagues can see what is going on.

```{r include=FALSE}
rm(p_grid)
rm(prior)
rm(prob_data)
rm(posterior)
rm(samples)
```

# Chapter 3

```{r}
## R code 3.1
Pr_Positive_Vampire <- 0.95 # true positive
Pr_Positive_Mortal <- 0.01  # false positive
Pr_Vampire <- 0.001  # prevalance 

# (probability of true positive * relative occurance true positive) + (probability false positive * relative occurance of false positive)
# i.e. average probability of data
Pr_Positive <- Pr_Positive_Vampire * Pr_Vampire +  # (probability of true positive * relative occurance true positive)
               Pr_Positive_Mortal * ( 1 - Pr_Vampire )  # (probability false positive * relative occurance of false positive)
( Pr_Vampire_Positive <- Pr_Positive_Vampire*Pr_Vampire / Pr_Positive )
```

---

```{r}
## R code 3.2
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)
```

```{r}
## R code 3.3
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

```{r}
## R code 3.5
dens( samples )
```


```{r}
dens( sample( p_grid , prob=posterior , size=1e3 , replace=TRUE ) )
dens( sample( p_grid , prob=posterior , size=1e4 , replace=TRUE ) )
dens( sample( p_grid , prob=posterior , size=1e5 , replace=TRUE ) )
```

```{r}
## R code 3.6
# add up posterior probability where p < 0.5
sum( posterior[ p_grid < 0.5 ] )
```

```{r}
## R code 3.7
sum( samples < 0.5 ) / 1e4
```

```{r}
## R code 3.8
sum( samples > 0.5 & samples < 0.75 ) / 1e4
```


```{r}
## R code 3.9
quantile( samples , 0.8 )
```


```{r}
## R code 3.10
quantile( samples , c( 0.1 , 0.9 ) )
```


```{r}
## R code 3.11
set.seed(1)
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep(1,1000)
likelihood <- dbinom( 3 , size=3 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
samples <- sample( p_grid , size=1e4 , replace=TRUE , prob=posterior )
```

```{r}
plot(p_grid, posterior)
```

```{r}
## R code 3.12
PI( samples , prob=0.5 )
```


```{r}
## R code 3.13
HPDI( samples , prob=0.5 )
```


```{r}
## R code 3.14
p_grid[ which.max(posterior) ]
```

```{r}
## R code 3.15
chainmode( samples , adj=0.01 )
```

```{r}
## R code 3.16
mean( samples )
median( samples )
```

---

Loss Functions, here is our posterior.

```{r}
plot(p_grid, posterior)
```

Absolute loss, pg 60

```{r}
## R code 3.17
sum( posterior*abs( 0.5 - p_grid ) )
```

```{r}
sum( posterior*abs( 0.8 - p_grid ) )
```

`posterior*abs( d - p_grid )` takes a particular guess at the "TRUE" value of p (the proportion of water in this example), and calculates the distance from our guess `d` to each of the possible points `p` which is defined by `p_grid`.

Now that we have the relative distance away from our guess for each possible value of p, we multiply that by posterior which gives higher numbers for more plausible values.

So the further a particular guess is from a particular point, the higher the distance and the more it will be penalized relative to the plausibility of that p.

Let's say our best guess is `0.5`. We'll calculate a single value in the calculation above, let's use `p=1` which corresponds with the mode of the graph. 

We are a distance of 0.5, and that posterior is going to penalize us the most since it has the highest plausiability.

```{r}
abs(0.5 - p_grid[1000]) 
posterior[1000]
posterior[1000] * abs(0.5 - p_grid[1000])
```

Now if our guess was `0.8` we wouldn't get penalized as much at that same point, because we are closer.

```{r}
abs(0.8 - p_grid[1000]) 
posterior[1000]
posterior[1000] * abs(0.8 - p_grid[1000])
```

Or a guess of `1.0`

Now we aren't penalized at all because we aren't any distance away.

```{r}
abs(1 - p_grid[1000]) 
posterior[1000]
posterior[1000] * abs(1 - p_grid[1000])
```

But this function does not optimize for the mode, because you are still calculating all the penalties across the entire space of plausible values (p_grid), for each guess.

So while a guess of 0.8 gives you a higher penalty at that specific value of 1, it turns out that summing across all the penalties will produce a lower overall penalty (for 0.8) than if you were to guess 1.0

Instead, we'll make a guess at each possible value of p (probability of water), which is defined by p_grid. At each value, we'll calculate our loss.

```{r}
## R code 3.18
loss <- sapply( p_grid , function(d) sum( posterior*abs( d - p_grid ) ) )
```

```{r}
## R code 3.19
p_grid[ which.min(loss) ]
```

```{r}
plot(p_grid, loss)
abline(v = p_grid[ which.min(loss) ], col="red", lwd=3, lty=2)
```

```{r}
## R code 3.20
dbinom( 0:2 , size=2 , prob=0.7 )
```


```{r}
## R code 3.21
rbinom( 1 , size=2 , prob=0.7 )
```


```{r}
## R code 3.22
rbinom( 10 , size=2 , prob=0.7 )
```


```{r}
## R code 3.23
dummy_w <- rbinom( 1e5 , size=2 , prob=0.7 )
table(dummy_w)/1e5
```


```{r}
## R code 3.24
dummy_w <- rbinom( 1e5 , size=9 , prob=0.7 )
simplehist( dummy_w , xlab="dummy water count" )
```


```{r}
## R code 3.25
w <- rbinom( 1e4 , size=9 , prob=0.6 )
w[1:20]
```

```{r}
hist(w)
```

```{r}
## R code 3.27
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e5 , replace=TRUE )
```

```{r}
## R code 3.26
w <- rbinom( 1e5 , size=9 , prob=samples )
length(w)
samples[1:20]
w[1:20]
```

So, we calculated the posterior probability based on our sample of observing 6 waters in 9 tosses.

We then sampled our posterior distribution, which gives us a list of possible proportions of water with relative frequency corresponding to posterior probability.

We then feed this into `rbinom` which produces random generation of how many times you'd expect to see water in 9 tosses, and uses the sampled proportions from our posterior distribution.

So we are propagating the uncertainty of the posterior into `rbinom` which is why the histogram below is more spread out (less certain) than the histogram above. The histogram above only shows the expected distribution if the true proportion of water is 0.

```{r}
hist(w)
```

```{r}
## R code 3.28
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)
```

```{r}
## R code 3.29
library(rethinking)
data(homeworkch3)
```

```{r}
## R code 3.30
sum(birth1) + sum(birth2)
```

```{r include=FALSE}
rm(Pr_Positive_Vampire)
rm(Pr_Positive_Mortal)
rm(Pr_Vampire)
rm(Pr_Positive)
rm(Pr_Vampire_Positive)
rm(p_grid)
rm(prior)
rm(prob_data)
rm(posterior)
rm(samples)
```

# Chapter 4

```{r}
## R code 4.1
pos <- replicate( 1000 , sum( runif(16,-1,1) ) )
```

```{r}
## R code 4.2
prod( 1 + runif(12,0,0.1) )
```

```{r}
## R code 4.3
growth <- replicate( 10000 , prod( 1 + runif(12,0,0.1) ) )
growth[1:20]
dens( growth , norm.comp=TRUE )
```

```{r}
## R code 4.4
big <- replicate( 10000 , prod( 1 + runif(12,0,0.5) ) )
small <- replicate( 10000 , prod( 1 + runif(12,0,0.01) ) )
```

```{r}
## R code 4.5
log.big <- replicate( 10000 , prod(1 + runif(12,0,0.5)) )
dens( log.big , norm.comp=TRUE )
```

```{r}
log.big.log <- replicate( 10000 , log(prod(1 + runif(12,0,0.5))) )
dens( log.big.log , norm.comp=TRUE )
```

> Adding logs is equivalent to multiplying the original numbers.

```{r}
exp(log(4) + log(13))
4*13
```


```{r}
## R code 4.6
w <- 6; n <- 9;
p_grid <- seq(from=0,to=1,length.out=100)
posterior <- dbinom(w,n,p_grid)*dunif(p_grid,0,1)
posterior <- posterior/sum(posterior)
```

```{r}
## R code 4.7
library(rethinking)
data(Howell1)
d <- Howell1
```

```{r}
## R code 4.8
str( d )
```

```{r}
## R code 4.9
precis( d )
```

```{r}
## R code 4.10
d$height
```

```{r}
## R code 4.11
d2 <- d[ d$age >= 18 , ]
```

```{r}
hist(d2$height)
dens(d2$height)
```


```{r}
## R code 4.12
curve( dnorm( x , 178 , 20 ) , from=100 , to=250 )
```

```{r}
## R code 4.13
curve( dunif( x , 0 , 50 ) , from=-10 , to=60 )
```

```{r}
## R code 4.14
sample_mu <- rnorm( 1e4 , 178 , 20 )
sample_sigma <- runif( 1e4 , 0 , 50 )
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )
```

```{r}
## R code 4.15
sample_mu <- rnorm( 1e4 , 178 , 100 )
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )
```

```{r}
## R code 4.16
mu.list <- seq( from=150, to=160 , length.out=100 )
sigma.list <- seq( from=7 , to=9 , length.out=100 )
post <- expand.grid( mu=mu.list , sigma=sigma.list )  # all possible combinations of mean and sigma that we are considering
# for each possible combination of mean/sigma sum the log densities of all the heights
post$LL <- sapply( 1:nrow(post) , function(i) sum(dnorm( d2$height , post$mu[i] , post$sigma[i] , log=TRUE ) ) )

 
#dnorm( d2$height , post$mu[1] , post$sigma[1] , log=TRUE )
#sum(dnorm( d2$height , post$mu[1] , post$sigma[1] , log=TRUE ) )

# Log Likelihood + Log Prior + Log Sigma
# then we do exp(Log Likelihood + Log Prior + Log Sigma)
# which is the same as
# Likelihood * Prior * Sigma)
post$prod <- post$LL + dnorm( post$mu , 178 , 20 , TRUE ) +
    dunif( post$sigma , 0 , 50 , TRUE )
post$prob <- exp( post$prod - max(post$prod) )
```

```{r}
sum(post$prob)
```

```{r}
dnorm(x = c(-1, 0, 1))
```

```{r}
plot( post$mu , post$sigma )
```

```{r}
## R code 4.17
contour_xyz( post$mu , post$sigma , post$prob )
```

```{r}
## R code 4.18
image_xyz( post$mu , post$sigma , post$prob )
```

```{r}
## R code 4.19
sample.rows <- sample( 1:nrow(post) , size=1e4 , replace=TRUE , prob=post$prob )
sample.mu <- post$mu[ sample.rows ]
sample.sigma <- post$sigma[ sample.rows ]
```

```{r}
## R code 4.20
plot( sample.mu , sample.sigma , cex=0.5 , pch=16 , col=col.alpha(rangi2,0.1) )
```

```{r}
## R code 4.21
dens( sample.mu )
dens( sample.sigma )
```

```{r}
## R code 4.22
PI( sample.mu )
PI( sample.sigma )
```

```{r}
## R code 4.23
d3 <- sample( d2$height , size=20 )
```

```{r}
## R code 4.24
mu.list <- seq( from=150, to=170 , length.out=200 )
sigma.list <- seq( from=4 , to=20 , length.out=200 )
post2 <- expand.grid( mu=mu.list , sigma=sigma.list )
post2$LL <- sapply( 1:nrow(post2) , function(i)
    sum( dnorm( d3 , mean=post2$mu[i] , sd=post2$sigma[i] ,
    log=TRUE ) ) )
post2$prod <- post2$LL + dnorm( post2$mu , 178 , 20 , TRUE ) +
    dunif( post2$sigma , 0 , 50 , TRUE )
post2$prob <- exp( post2$prod - max(post2$prod) )
sample2.rows <- sample( 1:nrow(post2) , size=1e4 , replace=TRUE ,
    prob=post2$prob )
sample2.mu <- post2$mu[ sample2.rows ]
sample2.sigma <- post2$sigma[ sample2.rows ]
plot( sample2.mu , sample2.sigma , cex=0.5 ,
    col=col.alpha(rangi2,0.1) ,
    xlab="mu" , ylab="sigma" , pch=16 )
```

```{r}
## R code 4.25
dens( sample2.sigma , norm.comp=TRUE )
```

```{r}
## R code 4.26
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]
```

```{r}
## R code 4.27
flist <- alist(
    height ~ dnorm( mu , sigma ) ,
    mu ~ dnorm( 178 , 20 ) ,
    sigma ~ dunif( 0 , 50 )
)
flist
```

```{r}
## R code 4.28
m4.1 <- quap( flist , data=d2 )
```

```{r}
## R code 4.29
precis( m4.1 )
```

```{r}
mean(d2$height)
sd(d2$height)
```

```{r}
print(suppressWarnings(stan_glm(d2$height ~ 1, refresh=0)))
print(suppressWarnings(stan_glm(d2$height ~ 1, prior_intercept = normal(178, 20), refresh=0)))
```

```{r}
## R code 4.30
start <- list(
    mu=mean(d2$height),
    sigma=sd(d2$height)
)
m4.1 <- quap( flist , data=d2 , start=start )
```

```{r}
## R code 4.31
m4.2 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu ~ dnorm( 178 , 0.1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
precis( m4.2 )
```

```{r}
## R code 4.32
vcov( m4.1 )
```

```{r}
## R code 4.33
diag( vcov( m4.1 ) )
cov2cor( vcov( m4.1 ) )
```

```{r}
## R code 4.34
library(rethinking)
post <- extract.samples( m4.1 , n=1e4 )
head(post)
```

```{r}
## R code 4.35
precis(post)
```


```{r}
stan_glm_simulations <- as.matrix(stan_glm(d2$height ~ 1, refresh=0)) %>% as.data.frame()
stan_glm_simulations %>% head()
```

```{r}
precis(stan_glm_simulations)
```



```{r}
## R code 4.36
post <- MASS::mvrnorm( n=1e4 , mu=coef(m4.1) , Sigma=vcov(m4.1) )
```

```{r}
## R code 4.37
library(rethinking)
data(Howell1); d <- Howell1; d2 <- d[ d$age >= 18 , ]
plot( d2$height ~ d2$weight )
```

```{r}
## R code 4.38
set.seed(2971)
N <- 100                   # 100 lines
a <- rnorm( N , 178 , 20 )
b <- rnorm( N , 0 , 10 )
```

```{r}
## R code 4.39
plot( NULL , xlim=range(d2$weight) , ylim=c(-100,400) ,
    xlab="weight" , ylab="height" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
mtext( "b ~ dnorm(0,10)" )
xbar <- mean(d2$weight)
for ( i in 1:N ) curve( a[i] + b[i]*(x - xbar) ,
    from=min(d2$weight) , to=max(d2$weight) , add=TRUE ,
    col=col.alpha("black",0.2) )
```

```{r}
## R code 4.40
b <- rlnorm( 1e4 , 0 , 1 )
dens( b , xlim=c(0,5) , adj=0.1 )
```

```{r}
## R code 4.41
set.seed(2971)
N <- 100                   # 100 lines
a <- rnorm( N , 178 , 20 )
b <- rlnorm( N , 0 , 1 )
```

```{r}
## R code 4.39
plot( NULL , xlim=range(d2$weight) , ylim=c(-100,400) ,
    xlab="weight" , ylab="height" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
mtext( "b ~ dnorm(0,10)" )
xbar <- mean(d2$weight)
for ( i in 1:N ) curve( a[i] + b[i]*(x - xbar) ,
    from=min(d2$weight) , to=max(d2$weight) , add=TRUE ,
    col=col.alpha("black",0.2) )
```

```{r}
## R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1); d <- Howell1; d2 <- d[ d$age >= 18 , ]

# define the average weight, x-bar
xbar <- mean(d2$weight)

# fit model
m4.3 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - xbar ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )

m4.3_not_centered <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*weight  ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
```

```{r}
## R code 4.43
m4.3b <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + exp(log_b)*( weight - xbar ),
        a ~ dnorm( 178 , 20 ) ,
        log_b ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
```

```{r}
## R code 4.44
precis( m4.3 )
```

```{r}
stan_model <- d2 %>%
    mutate(weight_centered = weight - mean(weight)) %>%
    stan_glm(height ~ weight_centered, data=., refresh=0)

stan_model %>%
    print(2)
```

```{r}
## R code 4.45
round( vcov( m4.3 ) , 3 )
```

```{r}
## R code 4.46
plot( height ~ weight , data=d2 , col=rangi2 )
post <- extract.samples( m4.3 )
a_map <- mean(post$a)
b_map <- mean(post$b)
curve( a_map + b_map*(x - xbar) , add=TRUE )
```

```{r}
## R code 4.47
post <- extract.samples( m4.3 )
post[1:5,]
```

```{r}
## R code 4.48
N <- 10
dN <- d2[ 1:N , ]
mN <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - mean(weight) ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=dN )
```

```{r}
## R code 4.49
# extract 20 samples from the posterior
post <- extract.samples( mN , n=20 )

# display raw data and sample size
plot( dN$weight , dN$height ,
    xlim=range(d2$weight) , ylim=range(d2$height) ,
    col=rangi2 , xlab="weight" , ylab="height" )
mtext(concat("N = ",N))

# plot the lines, with transparency
for ( i in 1:20 )
    curve( post$a[i] + post$b[i]*(x-mean(dN$weight)) ,
        col=col.alpha("black",0.3) , add=TRUE )
```

```{r}
## R code 4.50
post <- extract.samples( m4.3 )
mu_at_50 <- post$a + post$b * ( 50 - xbar )
```

```{r}
## R code 4.51
dens( mu_at_50 , col=rangi2 , lwd=2 , xlab="mu|weight=50" )
```

```{r}
## R code 4.52
PI( mu_at_50 , prob=0.89 )
```

```{r}
## R code 4.53
mu <- link( m4.3 )
str(mu)
```

```{r}
## R code 4.54
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from=25 , to=70 , by=1 )

# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m4.3 , data=data.frame(weight=weight.seq) )
str(mu)
```

```{r}
## R code 4.55
# use type="n" to hide raw data
plot( height ~ weight , d2 , type="n" )

# loop over samples and plot each mu value
for ( i in 1:100 )
    points( weight.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.2) )
```

```{r}
posterior_linpred(stan_model, newdata = data.frame(weight_centered=50 - mean(d2$weight))) %>%
    quantile(c(.055, 0.945))
```

```{r}
PI(mu_at_50, prob = 0.89)
```


```{r}
set.seed(1)
posterior_line <- posterior_linpred(stan_model, newdata = data.frame(weight_centered=seq(30 - round(mean(d2$weight)),
                                                                                         60 - round(mean(d2$weight)),
                                                                                         1)))
dim(posterior_line)
```

```{r}
t(posterior_line) %>%
    as.data.frame() %>%
    mutate(weight = seq(30, 60, 1)) %>%
    pivot_longer(-weight) %>%
    ggplot(aes(x=weight, y=value)) +
    geom_point(alpha=0.2)
```
4K simulations corresponding to each fitted data point


```{r}
## R code 4.56
# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.57
# plot raw data
# fading out points to make line and interval more visible
plot( height ~ weight , data=d2 , col=col.alpha(rangi2,0.5) )

# plot the MAP line, aka the mean mu for each weight
lines( weight.seq , mu.mean )

# plot a shaded region for 89% PI
shade( mu.PI , weight.seq )
```

```{r}
## R code 4.58
post <- extract.samples(m4.3)
mu.link <- function(weight) post$a + post$b*( weight - xbar )
weight.seq <- seq( from=25 , to=70 , by=1 )
mu <- sapply( weight.seq , mu.link )
mu.mean <- apply( mu , 2 , mean )
mu.CI <- apply( mu , 2 , PI , prob=0.89 )
```





```{r}
## R code 4.59
sim.height <- sim( m4.3 , data=data.frame(weight=50) )
dim(sim.height)
```

```{r}
PI(sim.height, prob=0.89)
```

```{r}
posterior_predict(stan_model, newdata = data.frame(weight_centered=50 - mean(d2$weight))) %>%
    quantile(c(.055, 0.945))
```

```{r}
set.seed(1)
posterior_predictions <- posterior_predict(stan_model,
                                           newdata = data.frame(weight_centered=seq(30 - round(mean(d2$weight)),
                                                                                    60 - round(mean(d2$weight)),
                                                                                    1)))
dim(posterior_predictions)
```

```{r}
simulations <- t(posterior_predictions) %>%
    as.data.frame() %>%
    mutate(weight = seq(30, 60, 1)) %>%
    pivot_longer(-weight)

prediction_intervals <- simulations %>%
    group_by(weight) %>%
    summarise(`5.5%` = quantile(value, 0.055),
              `94.5%` = quantile(value, 0.945)) %>%
    ungroup()

simulations %>%
    ggplot(aes(x=weight, y=value)) +
    geom_ribbon(data = prediction_intervals,
                aes(x=weight, ymin=`5.5%`, ymax=`94.5%`, y=NULL), alpha=0.4) +
    geom_point(alpha=0.2) +
    geom_point(data = d2, aes(x=weight, y=height), color='red')
```

```{r}
## R code 4.60
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.61
# plot raw data
plot( height ~ weight , d2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )

# draw PI region for simulated heights
shade( height.PI , weight.seq )
```

```{r}
## R code 4.62
sim.height <- sim( m4.3 , data=list(weight=weight.seq) , n=1e4 )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.63
post <- extract.samples(m4.3)
weight.seq <- 25:70
sim.height <- sapply( weight.seq , function(weight)
    rnorm(
        n=nrow(post) ,
        mean=post$a + post$b*( weight - xbar ) ,
        sd=post$sigma ) )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.64
library(rethinking)
data(Howell1)
d <- Howell1
```

```{r}
## R code 4.65
d$weight_s <- ( d$weight - mean(d$weight) )/sd(d$weight)
d$weight_s2 <- d$weight_s^2
m4.5 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight_s + b2*weight_s2 ,
        a ~ dnorm( 178 , 20 ) ,
        b1 ~ dlnorm( 0 , 1 ) ,
        b2 ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d )
```

```{r}
## R code 4.66
precis( m4.5 )
```

```{r}
set.seed(1)
d %>%
    mutate(weight =(weight - mean(weight)) / sd(weight),
           weight_2 = weight^2) %>%
    stan_glm(height ~ weight + weight_2 , 
             prior=list(weight=normal(0, 1), weight_2=normal(0,1)) data = ., refresh=0) %>%
    tidy()
```


```{r}
## R code 4.67
weight.seq <- seq( from=-2.2 , to=2 , length.out=30 )
pred_dat <- list( weight_s=weight.seq , weight_s2=weight.seq^2 )
mu <- link( m4.5 , data=pred_dat )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )
sim.height <- sim( m4.5 , data=pred_dat )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.68
plot( height ~ weight_s , d , col=col.alpha(rangi2,0.5) )
lines( weight.seq , mu.mean )
shade( mu.PI , weight.seq )
shade( height.PI , weight.seq )
```

```{r}
## R code 4.69
d$weight_s3 <- d$weight_s^3
m4.6 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight_s + b2*weight_s2 + b3*weight_s3 ,
        a ~ dnorm( 178 , 20 ) ,
        b1 ~ dlnorm( 0 , 1 ) ,
        b2 ~ dnorm( 0 , 10 ) ,
        b3 ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d )
```

```{r}
## R code 4.70
plot( height ~ weight_s , d , col=col.alpha(rangi2,0.5) , xaxt="n" )
```

```{r}
## R code 4.71
at <- c(-2,-1,0,1,2)
labels <- at*sd(d$weight) + mean(d$weight)
axis( side=1 , at=at , labels=round(labels,1) )
```

```{r}
## R code 4.72
library(rethinking)
data(cherry_blossoms)
d <- cherry_blossoms
precis(d)
```

```{r}
## R code 4.73
d2 <- d[ complete.cases(d$doy) , ] # complete cases on doy
num_knots <- 15
knot_list <- quantile( d2$year , probs=seq(0,1,length.out=num_knots) )
```

```{r}
## R code 4.74
library(splines)
B <- bs(d2$year,
    knots=knot_list[-c(1,num_knots)] ,
    degree=15 , intercept=TRUE )
dim(B)
```

```{r}
head(B)
#B %>% View()
```

```{r}
## R code 4.75
plot( NULL , xlim=range(d2$year) , ylim=c(0,1) , xlab="year" , ylab="basis" )
for ( i in 1:ncol(B) ) lines( d2$year , B[,i] )
```

```{r}
## R code 4.76
m4.7 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + B %*% w ,
        a ~ dnorm(100,10),
        w ~ dnorm(0,10),
        sigma ~ dexp(1)
    ), data=list( D=d2$doy , B=B ) ,
    start=list( w=rep( 0 , ncol(B) ) ) )
```

```{r}
## R code 4.77
post <- extract.samples( m4.7 )
w <- apply( post$w , 2 , mean )
plot( NULL , xlim=range(d2$year) , ylim=c(-6,6) ,
    xlab="year" , ylab="basis * weight" )
for ( i in 1:ncol(B) ) lines( d2$year , w[i]*B[,i] )
```

```{r}
## R code 4.78
mu <- link( m4.7 )
mu_PI <- apply(mu,2,PI,0.97)
plot( d2$year , d2$doy , col=col.alpha(rangi2,0.3) , pch=16 )
shade( mu_PI , d2$year , col=col.alpha("black",0.5) )
```

```{r}
## R code 4.79
m4.7alt <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + sapply( 1:827 , function(i) sum( B[i,]*w ) ) ,
        a ~ dnorm(100,1),
        w ~ dnorm(0,10),
        sigma ~ dexp(1)
    ),
    data=list( D=d2$doy , B=B ) ,
    start=list( w=rep( 0 , ncol(B) ) ) )
```


# Chapter 5

```{r}
## R code 5.1
# load data and copy
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce
```

# standardize variables
d$D <- standardize( d$Divorce )
d$M <- standardize( d$Marriage )
d$A <- standardize( d$MedianAgeMarriage )
```


```{r}
## R code 5.2
sd( d$MedianAgeMarriage )

## R code 5.3
m5.1 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bA * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )

## R code 5.4
set.seed(10)
prior <- extract.prior( m5.1 )
mu <- link( m5.1 , post=prior , data=list( A=c(-2,2) ) )
plot( NULL , xlim=c(-2,2) , ylim=c(-2,2) )
for ( i in 1:50 ) lines( c(-2,2) , mu[i,] , col=col.alpha("black",0.4) )

## R code 5.5
# compute percentile interval of mean
A_seq <- seq( from=-3 , to=3.2 , length.out=30 )
mu <- link( m5.1 , data=list(A=A_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( D ~ A , data=d , col=rangi2 )
lines( A_seq , mu.mean , lwd=2 )
shade( mu.PI , A_seq )

## R code 5.6
m5.2 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM * M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )

## R code 5.7
library(dagitty)
dag5.1 <- dagitty( "dag{ A -> D; A -> M; M -> D }" )
coordinates(dag5.1) <- list( x=c(A=0,D=1,M=2) , y=c(A=0,D=1,M=0) )
drawdag( dag5.1 )

## R code 5.8
DMA_dag2 <- dagitty('dag{ D <- A -> M }')
impliedConditionalIndependencies( DMA_dag2 )

## R code 5.9
DMA_dag1 <- dagitty('dag{ D <- A -> M -> D }')
impliedConditionalIndependencies( DMA_dag1 )

## R code 5.10
m5.3 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis( m5.3 )

## R code 5.11
plot( coeftab(m5.1,m5.2,m5.3), par=c("bA","bM") )

## R code 5.12
N <- 50 # number of simulated States
age <- rnorm( N )        # sim A
mar <- rnorm( N , -age )  # sim A -> M
div <- rnorm( N , age )  # sim A -> D

## R code 5.13
m5.4 <- quap(
    alist(
        M ~ dnorm( mu , sigma ) ,
        mu <- a + bAM * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bAM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )

## R code 5.14
mu <- link(m5.4)
mu_mean <- apply( mu , 2 , mean )
mu_resid <- d$M - mu_mean

## R code 5.15
# call link without specifying new data
# so it uses original data
mu <- link( m5.3 )

# summarize samples across cases
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate observations
# again no new data, so uses original data
D_sim <- sim( m5.3 , n=1e4 )
D_PI <- apply( D_sim , 2 , PI )

## R code 5.16
plot( mu_mean ~ d$D , col=rangi2 , ylim=range(mu_PI) ,
    xlab="Observed divorce" , ylab="Predicted divorce" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d) ) lines( rep(d$D[i],2) , mu_PI[,i] , col=rangi2 )

## R code 5.17
identify( x=d$D , y=mu_mean , labels=d$Loc )

## R code 5.18
N <- 100                         # number of cases
x_real <- rnorm( N )             # x_real as Gaussian with mean 0 and stddev 1
x_spur <- rnorm( N , x_real )    # x_spur as Gaussian with mean=x_real
y <- rnorm( N , x_real )         # y as Gaussian with mean=x_real
d <- data.frame(y,x_real,x_spur) # bind all together in data frame

## R code 5.19
data(WaffleDivorce)
d <- list()
d$A <- standardize( WaffleDivorce$MedianAgeMarriage )
d$D <- standardize( WaffleDivorce$Divorce )
d$M <- standardize( WaffleDivorce$Marriage )

m5.3_A <- quap(
    alist(
      ## A -> D <- M
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 ),
      ## A -> M
        M ~ dnorm( mu_M , sigma_M ),
        mu_M <- aM + bAM*A,
        aM ~ dnorm( 0 , 0.2 ),
        bAM ~ dnorm( 0 , 0.5 ),
        sigma_M ~ dexp( 1 )
    ) , data = d )

## R code 5.20
A_seq <- seq( from=-2 , to=2 , length.out=30 )

## R code 5.21
# prep data
sim_dat <- data.frame( A=A_seq )

# simulate M and then D, using A_seq
s <- sim( m5.3_A , data=sim_dat , vars=c("M","D") )

## R code 5.22
plot( sim_dat$A , colMeans(s$D) , ylim=c(-2,2) , type="l" ,
    xlab="manipulated A" , ylab="counterfactual D"  )
shade( apply(s$D,2,PI) , sim_dat$A )
mtext( "Total counterfactual effect of A on D" )

## R code 5.23
# new data frame, standardized to mean 26.1 and std dev 1.24
sim2_dat <- data.frame( A=(c(20,30)-26.1)/1.24 )
s2 <- sim( m5.3_A , data=sim2_dat , vars=c("M","D") )
mean( s2$D[,2] - s2$D[,1] )

## R code 5.24
sim_dat <- data.frame( M=seq(from=-2,to=2,length.out=30) , A=0 )
s <- sim( m5.3_A , data=sim_dat , vars="D" )

plot( sim_dat$M , colMeans(s) , ylim=c(-2,2) , type="l" ,
    xlab="manipulated M" , ylab="counterfactual D"  )
shade( apply(s,2,PI) , sim_dat$M )
mtext( "Total counterfactual effect of M on D" )

## R code 5.25
A_seq <- seq( from=-2 , to=2 , length.out=30 )

## R code 5.26
post <- extract.samples( m5.3_A )
M_sim <- with( post , sapply( 1:30 ,
    function(i) rnorm( 1e3 , aM + bAM*A_seq[i] , sigma_M ) ) )

## R code 5.27
D_sim <- with( post , sapply( 1:30 ,
    function(i) rnorm( 1e3 , a + bA*A_seq[i] + bM*M_sim[,i] , sigma ) ) )

## R code 5.28
library(rethinking)
data(milk)
d <- milk
str(d)

## R code 5.29
d$K <- standardize( d$kcal.per.g )
d$N <- standardize( d$neocortex.perc )
d$M <- standardize( log(d$mass) )

## R code 5.30
m5.5_draft <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N ,
        a ~ dnorm( 0 , 1 ) ,
        bN ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) , data=d )

## R code 5.31
d$neocortex.perc

## R code 5.32
dcc <- d[ complete.cases(d$K,d$N,d$M) , ]

## R code 5.33
m5.5_draft <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N ,
        a ~ dnorm( 0 , 1 ) ,
        bN ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )

## R code 5.34
prior <- extract.prior( m5.5_draft )
xseq <- c(-2,2)
mu <- link( m5.5_draft , post=prior , data=list(N=xseq) )
plot( NULL , xlim=xseq , ylim=xseq )
for ( i in 1:50 ) lines( xseq , mu[i,] , col=col.alpha("black",0.3) )

## R code 5.35
m5.5 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )

## R code 5.36
precis( m5.5 )

## R code 5.37
xseq <- seq( from=min(dcc$N)-0.15 , to=max(dcc$N)+0.15 , length.out=30 )
mu <- link( m5.5 , data=list(N=xseq) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( K ~ N , data=dcc )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )

## R code 5.38
m5.6 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
precis(m5.6)

## R code 5.39
m5.7 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
precis(m5.7)

## R code 5.40
plot( coeftab( m5.5 , m5.6 , m5.7 ) , pars=c("bM","bN") )

## R code 5.41
xseq <- seq( from=min(dcc$M)-0.15 , to=max(dcc$M)+0.15 , length.out=30 )
mu <- link( m5.7 , data=data.frame( M=xseq , N=0 ) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( NULL , xlim=range(dcc$M) , ylim=range(dcc$K) )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )

## R code 5.42
# M -> K <- N
# M -> N
n <- 100
M <- rnorm( n )
N <- rnorm( n , M )
K <- rnorm( n , N - M )
d_sim <- data.frame(K=K,N=N,M=M)

## R code 5.43
# M -> K <- N
# N -> M
n <- 100
N <- rnorm( n )
M <- rnorm( n , N )
K <- rnorm( n , N - M )
d_sim2 <- data.frame(K=K,N=N,M=M)

# M -> K <- N
# M <- U -> N
n <- 100
U <- rnorm( n )
N <- rnorm( n , U )
M <- rnorm( n , U )
K <- rnorm( n , N - M )
d_sim3 <- data.frame(K=K,N=N,M=M)

## R code 5.44
dag5.7 <- dagitty( "dag{
    M -> K <- N
    M -> N }" )
coordinates(dag5.7) <- list( x=c(M=0,K=1,N=2) , y=c(M=0.5,K=1,N=0.5) )
MElist <- equivalentDAGs(dag5.7)

## R code 5.45
data(Howell1)
d <- Howell1
str(d)

## R code 5.46
mu_female <- rnorm(1e4,178,20)
mu_male <- rnorm(1e4,178,20) + rnorm(1e4,0,10)
precis( data.frame( mu_female , mu_male ) )

## R code 5.47
d$sex <- ifelse( d$male==1 , 2 , 1 )
str( d$sex )

## R code 5.48
m5.8 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a[sex] ,
        a[sex] ~ dnorm( 178 , 20 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d )
precis( m5.8 , depth=2 )

## R code 5.49
post <- extract.samples(m5.8)
post$diff_fm <- post$a[,1] - post$a[,2]
precis( post , depth=2 )

## R code 5.50
data(milk)
d <- milk
levels(d$clade)

## R code 5.51
d$clade_id <- as.integer( d$clade )

## R code 5.52
d$K <- standardize( d$kcal.per.g )
m5.9 <- quap(
    alist(
        K ~ dnorm( mu , sigma ),
        mu <- a[clade_id],
        a[clade_id] ~ dnorm( 0 , 0.5 ),
        sigma ~ dexp( 1 )
    ) , data=d )
labels <- paste( "a[" , 1:4 , "]:" , levels(d$clade) , sep="" )
plot( precis( m5.9 , depth=2 , pars="a" ) , labels=labels ,
    xlab="expected kcal (std)" )

## R code 5.53
set.seed(63)
d$house <- sample( rep(1:4,each=8) , size=nrow(d) )

## R code 5.54
m5.10 <- quap(
    alist(
        K ~ dnorm( mu , sigma ),
        mu <- a[clade_id] + h[house],
        a[clade_id] ~ dnorm( 0 , 0.5 ),
        h[house] ~ dnorm( 0 , 0.5 ),
        sigma ~ dexp( 1 )
    ) , data=d )
```

